### 2017-03-03 thesis/discussion/discussion.tex
tmp commit

\subsection{\gls{blca} data set and the obesity metagenes}
\label{sub:blca_and_obesity_metagenes}

@@

Ideally, multiple measurements that are indicative of obesity should be made in order to determine the obesity status of the patients to make a more accurate classification.
% Taking these extra measurements could be the difference between 
Additional measurements such as blood lipid profiles and insulin levels may also help understand the underlying biological state in which the patients are in, and would greatly help focus on the genes that are truly affecting the patients.
However, these measurements will be difficult to carry out in a systematic fashion in a clinical setting, and to collate existing patient data with these measurements to make a data set that is large enough for the analyses done in this project may be impossible.
Nevertheless, even the measurement of waist circumference will significantly help describe the obesity phenotype more accurately in future investigations.

@@

Second limitation was the quality of the data sets in which this project have used.
All of the data sets used in this project were from an online and publicly available source, and the types of data sets used were either microarray or \gls{rnaseq} data.

First problem -- different studies generated different data sets in a different way/method

second problem -- micorarry vs. rna seq data

third problem -- pure quality of the data

@@

The final limitation to consider was the quality of the genetic signatures used in this project.
This limitation is partly related to the previous section, as all of the genetic signatures were derived from the data sets used by other groups.

>>>>

\subsection{\Gls{blca} data set and the obesity metagenes}
\label{sub:blca_and_obesity_metagenes}

@@

Ideally, multiple measurements that are indicative of obesity should be made in order to determine the obesity status of the patients to make a more accurate classification.
Additional measurements such as blood lipid profiles and insulin levels may also help understand the underlying biological state in which the patients are in, and would greatly help focus on the genes that are truly affecting the patients.
However, these measurements will be difficult to carry out in a systematic fashion in a clinical setting, and to collate existing patient data with these measurements to make a data set that is large enough for the analyses done in this project may be impossible.
Nevertheless, even the measurement of waist circumference in addition to \gls{bmi} may significantly help describe the obesity phenotype more accurately in future investigations.

@@

Second limitation was the quality of the data sets in which this project have used.
All of the data sets used in this project were from an online and publicly available source, and the types of data sets used were either microarray or \gls{rnaseq} data.
First potential problem with using an open source data from other research groups is that the data is generated by multiple different groups with different technologies and experimental protocols.
Since different laboratories have their own way to carry out certain experiments, the sequence data from each of the studies used in this project may have been produced and processed in a completely different manner, not to mention about the use of different patient cohorts in each experiment.

For example, in the \citet{Creighton2012} study, the patient breast tumour biopsies were ``trimmed such that all the samples had $\geq$ 70\% tumour cells'' in the samples.
On the other hand, \citet{Fuentes-Mattei2014} only mentioned that they used breast tumour biopsies from the patient with no comment about the quality or quantity of the biopsies.
However, they took one third of the tumours from the mouse models to analyse the transcriptome of the mouse tumours, so perhaps Fuentes-Mattei \textit{et al.} used similar approach with the human breast tumour samples.
Without a doubt, the difference in how the samples are handled will affect the quality of the data, making it challenging to compare different data sets and it may even be possible for the samples within the data set to be incomparable.

% TODO: move some of this stuff into the false positive section
In fact, a couple of observations were made in this project that suggested the Creighton \textit{et al.} data set was of poor quality.
First in \cref{sub:identification_of_obesity_associated_genetic_signatures}, there were many \glspl{deg} identified from the gene expression analyses when obesity associated genetic signatures were sought for in the Creighton \textit{et al.} data set.
This suggested that there was a significant number of genes that were not related to obesity (false positives) in the genetic signatures.
Furthermore, the fact that about a third of the genes in any one of the obesity signatures were ``unique'' supported the apparent abundance of false positives in the signatures (\cref{fig:venn1,fig:venn2}).
From these results, it was likely that the obesity signatures that resulted from the Creighton \textit{et al.} data set were full of false positives due to the poor quality of the data.

Although the underlying quality of the data provided by these studies could not be refined in any way, best efforts were made to make sure the data sets were free of other factors that may affect the results.
% Although the underlying quality of the data provided by these studies could not be refined in any way, best efforts were made to make sure the results from these data sets were reliable.
This leads to the second problem: the different sequencing technologies used in the studies.
All of Creighton \textit{et al.}, Fuentes-Mattei \textit{et al.}, Print \textit{et al.} and Gatza \textit{et al.} data sets were analysed with the microarray technology, whereas all of the \gls{icgc} cancer data sets were sequenced with \gls{ngs} technology.
As mentioned in \cref{sub:data_normalisation}, the two technologies use fundamentally different principals to output sequence information of the sample; microarray uses light intensity and \gls{ngs} uses count data.

In order to make these two different types of data comparable to one another, normalisation methods from the \textit{limma} package were used (see \cref{sub:data_normalisation}).
This enabled the standard \textit{limma} analysis pipelines, normally used for the analysis of microarray data, to be used by the \gls{icgc} cancer \gls{rnaseq} data sets.
In this way, the results from the two types of data sets were fairly reliable and consistent as the same analysis methods were applied to the data sets.
In addition to the difference in the sequencing technologies, the Gatza \textit{et al.} data set comprised of multiple different data sets, each from different sources.
Since individual data set had their own experimental differences (``batch effect''), it was important to correct for these differences when multiple data sets were combined into one (see \cref{sub:batch_correction}).
In this project, batch corrections were made in the Gatza \textit{et al.} data set and the combined \gls{icgc} cancer data sets (used in \cref{sec:pathways_enriched_in_icgc_data_sets}) to eliminate the batch effects.

It was evident from some of the results that the Creighton \textit{et al.} data set was of poor quality, and the validity of the other data sets (and therefore the results generated from these data sets) were also debatable, despite all of the attempts made to control for other factors that may have affected the results.

@@

The final limitation to consider was the quality of the genetic signatures used in this project.
This limitation is partly related to \cref{sub:quality_of_the_data}, as  the identification of the  genetic signatures depended  on the quality of the data as well as the definition  of the signature in the first instance (for example, obesity or Akt overexpression).

Firstly, the obesity signatures identified by both Creighton \textit{et al.} and Fuentes-Mattei \textit{et al.} relied on the definition of obesity based on the sample \gls{bmi} values.
As discussed in \cref{sub:discussion_definition_of_obesity}, the use of \gls{bmi} may not have been the best measurement for central adiposity.
In fact, both Creighton \textit{et al.} and Fuentes-Mattei \textit{et al.} signatures associated with different pathway associated signatures in \cref{sec:pathway_associated_metagenes_and_obesity_associated_metagenes,sec:prediction_of_obesity_associated_metagene_with_pathway_associate_metagene}, which suggested that obesity as defined by \gls{bmi} was not enough to reveal the underlying biological characteristic of obesity, and hence showed association with different pathway signatures.
Furthermore, as mentioned in \cref{sub:quality_of_the_data}, it was likely that the Creighton \textit{et al.} data set, and perhaps other data sets as well, was of poor quality and may have affected the ability of the signatures as a marker for obesity in cancer.

Secondly for the pathway associated genetic signatures,
